\section*{Declaration}
Hereby I declare that this thesis is my original authorial work, which I have worked out on my own. All
sources, references, and literature used or excerpted during elaboration of this work are properly cited and
listed in complete reference to the due source.
\\
During the preparation of this thesis, the following AI tools were used:

\begin{itemize}
    \item \textbf{ChatGPT}: Used for debugging and making small code corrections, as well as LaTeX formatting.
    \item \textbf{V0}: Used for the initial styling configuration.
\end{itemize}

I declare that they were used in accordance with the principles of academic integrity.

I checked the content and took full responsibility for it.

\vfill\noindent
\textbf{Thesis Advisor:} Mgr. Luděk Bártek, Ph.D
\cleardoublepage

\section*{Abstract}
This bachelor's thesis implements a WEB-based music streaming platform
with additional social features - live comment sections for songs and playlists,
discussion forums, additional possibilities for interaction with the followed users etc.

A study is made beforehand in order to determine what users expect from the application;
comparison and exploration of different existing platforms is presented in order
to give a better insight into the market of similar solutions.

The thesis leverages existing backend and frontend frameworks, such as Django and React, for the
actual handling of the underlying data, logical processes and the interface of the platform.
In addition, modern audio representation and streaming technologies,
such as MPEG-DASH and HLS are used.
In order for the application to feel responsive, Server Sent Events are added to provide
two-way communication between the client and the server - this ensures that
individual interactions are always synchronized with other users and instances of the program.


\subsection*{Keywords}
Audio Streaming, Python, Django, React, MPEG-DASH, SSE
